---
title: "Data Visualization 1"
author: "Youssra Yemmas"
date: "2023-09-28"
output: html_document
---
# GGPLOT- grammar of graphics. Organize data set into a data fram and organize plot froms the variables in the data frame. Basic graph components should have datset, asethetic mappings, geoms.
# Advanced graph components- facets, scales, statistics.
# A graph is built by combining these components. Components are consistent across graph types- scatterplots, bar graphs, density plots, ridges plots etc
```{r setup, include=FALSE}
library(tidyverse)
library(rnoaa)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```


```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
select(name, id, everything())
  
weather_df
```


## Including Plots


```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()
```

# Pipes and Stuff
```{r}
weather_df |>
  filter(name == "CentralPark_NY") |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
  
ggp_nyc_weather =
  weather_df |>
  filter(name == "CentralPark_NY") |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

## Fancy Plot

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point() + # you can move the aes(color = name) from the ggplot to change where you define your aesthetic, can also add an alpha= 0.3 to make the points less prominents
  geom_smooth(se = FALSE) # fits a smooth line through the three name groups # se false gets rid of the error bars
```

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3) +
  geom_smooth() +
  facet_grid(. ~ name)
```

```{r}
##ggplot(weather_df, aes(x = date, y = tmax, color = mean)) + 
  ##geom_point(aes(size = prcp), alpha = .3)
  ##geom_smooth()
  ##facet_grid(. ~ name)
  
```

```{r}
##weather_df |>
##  filter(name != "CentralPark_NY") |>
 ## rename(Data = date) |>
 ## ggplot(aes(x = date, y = tmax, color = name)) +
  ##geom_point(alpha = .7, size = .5)
```


```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()
```


```{r}
weather_df |>
  filter(name == "Molokai_HI") |>
  ggplot(aes(x = date, y = tmax)) +
  geom_line(alpha = .5) +
  geom_point(size = .5) 
```

## Univariate plotting 
```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) +
  geom_density(alpha = .3, adjust = 2)
  ## geom_histogram(position = "dodge")
```

## Compare univariate distrib using boxplots
```{r}
ggplot(weather_df, aes(y = tmax, x = name)) +
  geom_boxplot()
```

## Violin plots which is a good comparison plot
```{r}
ggplot(weather_df, aes(y = tmax, x = name)) +
  geom_violin()
```

## Ridge Plot
```{r}
##ggplot(weather_df, aes(x = tmax, y = name)) +
 ## geom_density_ridges()
```

# Saving and embedding plots
```{r}
ggp_weather = 
  weather_df |>
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()

ggp_weather
ggsave("ggp_weather.pdf", ggp_weather)
```

```{r, fig.width= 6, fig.height= 6, fig.asp=.6}
ggp_weather
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
