---
title: "Data Visualization 1"
author: "Youssra Yemmas"
date: "2023-09-28"
output: html_document
---
# GGPLOT- grammar of graphics. Organize data set into a data fram and organize plot froms the variables in the data frame. Basic graph components should have datset, asethetic mappings, geoms.
# Advanced graph components- facets, scales, statistics.
# A graph is built by combining these components. Components are consistent across graph types- scatterplots, bar graphs, density plots, ridges plots etc
```{r setup, include=FALSE}
library(tidyverse)
library(rnoaa)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```


```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
select(name, id, everything())
  
weather_df
```


## Same plot as last times 

```{r}
weather_df |>
  ggplot(aes(x = tmin, y= tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Min Daily temp Degrees C",
    y = "Max daily temp",
    color = "Location",
    caption = "max vs min daily temp in three locations from 2021 and 20233; data from RNOA"
  )
```


## Scales- try to modify the default mapping
```{r}
weather_df |>
  filter(tmax >= 20, tmax <= 30) |>
  ggplot(aes(x = tmin, y= tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Min Daily temp Degrees C",
    y = "Max daily temp",
    color = "Location",
    caption = "max vs min daily temp in three locations from 2021 and 2022; data from RNOA"
  ) +
  scale_x_continuous(
    breaks = c(-15, 0, 15),
    labels = c("-15 C", "0", "15")
  ) +
  scale_y_continuous(
    position = "right"
  )
```


what about colors ...

```{r}
weather_df |>
  ggplot(aes(x = tmin, y= tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Min Daily temp Degrees C",
    y = "Max daily temp",
    color = "Location",
    caption = "max vs min daily temp in three locations from 2021 and 2022; data from RNOA"
  ) + 
  viridis::scale_color_viridis(discrete = TRUE)
```


## Themes
```{r}
weather_df |>
  ggplot(aes(x = tmin, y= tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Min Daily temp Degrees C",
    y = "Max daily temp",
    color = "Location",
    caption = "max vs min daily temp in three locations from 2021 and 2022; data from RNOA"
  ) + 
  viridis::scale_color_viridis(discrete = TRUE) + 
  theme_minimal() +
  theme(legend.position = "bottom")
```



## data argument ...


```{r}
weather_df |>
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() + 
  geom_smooth()



nyc_weather_df =
  weather_df |>
  filter(name == "CentralPark_NY")

hawaii_weather_df = 
  weather_df |>
  filter(name == "Molokai_HI")

ggplot(nyc_weather_df, aes(x = dtae, y = tmax, color = name)) +
  geom_point() +
  geom_line(data = hawaii_weather_df)
```


## 'patchwork'


```{r}
weather_df |>
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  facet_grid(. ~ name)
```

```{r}
ggp_temp_scatter =
weather_df |>
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(alpha = .5)

ggp_prcp_density =
  weather_df |>
  filter(prcp > 25) |>
  ggplot(aes(x = prcp, fill = name) +
  geom_density(alpha = .5) +
  theme(legend.position = "bottom")
  
ggp_tmax_date = 
  weather_df |>
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE) + 
  theme(legend.position = "bottom")
```



##data manipulation

```{r}
weather_df |>
  mutate(
    name = fct_relevel(name, c("Molokai_HI", "CentralPark_NY", "Waterhole_WA"))
  ) |>
  ggplot(aes(x = name, y = tmax)) + 
  geom_boxplot()

weather_df |>
  mutate(
    name = fct_reorder(name, tmax)
  ) |>
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin()

```

